---
title: "EDS 231: Assignment 2"
author: "Scout Leonard"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: pdf_document
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, 
                      fig.height = 5, 
                      echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

# Objective

# Load Libararies

```{r}
# load packages
packages=c("tidyr",
           "lubridate",
           "pdftools",
           "pdftools",
           "tidytext",
           "here",
           "LexisNexisTools",
           "sentimentr",
           "readr",
           "textdata",
           "dplyr",
           "stringr",
           "janitor",
           "ggplot2")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}
```

# Part 0

**Using the “IPCC” Nexis Uni data set from the class presentation and the pseudo code we discussed, recreate Figure 1A from Froelich et al. (Date x # of 1) positive, 2) negative, 3) neutral headlines):**

```{r}
#to follow along with this example, download this .docx to your working directory: 
#https://github.com/MaRo406/EDS_231-text-sentiment/blob/main/nexis_dat/Nexis_IPCC_Results.docx
ipcc_files <- list.files(pattern = ".docx", path = here("data"),
                       full.names = TRUE, recursive = TRUE, ignore.case = TRUE)

ipcc_dat <- lnt_read(ipcc_files) #Object of class 'LNT output'

ipcc_meta_df <- ipcc_dat@meta
ipcc_articles_df <- ipcc_dat@articles
ipcc_paragraphs_df <- ipcc_dat@paragraphs

ipcc_dat2<- data_frame(element_id = seq(1:length(ipcc_meta_df$Headline)), 
                       Date = ipcc_meta_df$Date, 
                       Headline = ipcc_meta_df$Headline)

#May be of use for assignment: using the full text from the articles
ipcc_paragraphs_dat <- data_frame(element_id = ipcc_paragraphs_df$Art_ID, 
                             Text  = ipcc_paragraphs_df$Paragraph)

ipcc_dat3 <- inner_join(ipcc_dat2,
                        ipcc_paragraphs_dat, 
                        by = "element_id")
```

Use the Bing sentiment analysis lexicon.

```{r}
bing_sent <- get_sentiments('bing') #grab the bing sentiment lexicon from tidytext
head(bing_sent, n = 20)
```

```{r}
ipcc_text_words <- ipcc_dat3 %>%
  unnest_tokens(output = word, 
                input = Text, 
                token = 'words')

 ipcc_sent_words <- ipcc_text_words %>% #break text into individual words
  anti_join(stop_words, by = 'word') %>% #returns only the rows without stop words
  inner_join(bing_sent, by = 'word') %>% 
   clean_names()
 
```


```{r}
ipcc_sent_plot <- ipcc_sent_words %>% 
  count(sentiment, date)

ggplot(data = ipcc_sent_plot, aes(x = date, y = n)) +
         geom_line(aes(color = sentiment)) +
  theme_minimal() +
  labs(y = "Developed Media Sentiment (no. headlines)",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m-%Y") +
  scale_color_manual(values = c("blue", "red"))
```

# Part 1

**[Access the Nexis Uni database through the UCSB library] (https://www.library.ucsb.edu/research/db/211)**

\noindent Got it! 

\newpage

# Part 2

**Choose a key search term or terms to define a set of articles.**

\noindent Done! I chose the term, "school lunch." My MEDS cohort knows I love talking about the USDA National School Lunch Program... 

\newpage

# Part 3

**Use your search term along with appropriate filters to obtain and download a batch of at least 100 full text search results (.docx)..**

\noindent Sweet! All downloaded. 

\newpage

# Part 4

**Read your Nexis article document into RStudio.**

\noindent Now for some coding...

```{r}
lunch_files <- list.files(pattern = ".docx", 
                          path = here("data"),
                          full.names = TRUE, 
                          recursive = TRUE, 
                          ignore.case = TRUE)

lunch_dat <- lnt_read(lunch_files) #Object of class 'LNT output'

lunch_meta_df <- lunch_dat@meta
lunch_articles_df <- lunch_dat@articles
lunch_paragraphs_df <- lunch_dat@paragraphs

lunch_dat2<- data.frame(element_id = seq(1:length(lunch_meta_df$Headline)), 
                        Date = lunch_meta_df$Date, 
                        Headline = lunch_meta_df$Headline)
```

\newpage

# Part 5

**This time use the full text of the articles for the analysis. First clean any artifacts of the data collection process (hint: this type of thing should be removed: “Apr 04, 2022( Biofuels Digest: http://www.biofuelsdigest.com/ Delivered by Newstex”)).**

```{r}
lunch_paragraphs_dat <- data.frame(element_id = lunch_paragraphs_df$Art_ID, 
                             Text = lunch_paragraphs_df$Paragraph)

lunch_dat3 <- inner_join(lunch_dat2,
                   lunch_paragraphs_dat, 
                   by = "element_id") %>% 
  clean_names()

#unnest to word-level tokens, remove stop words, and join sentiment words
 lunch_text_words <- lunch_dat3  %>%
  unnest_tokens(output = word, 
                input = text, 
                token = 'words')
 
lunch_text_words <- lunch_text_words %>%
  anti_join(stop_words) #removes the stop words 
 
 #remove numbers 
lunch_text_words_list <- str_remove_all(lunch_text_words$word, "[:digit:]")
#remove website strings
lunch_text_words_list <- str_remove_all(lunch_text_words_list, "https")

#remove the empty strings
tib <-subset(lunch_text_words_list != "")

#reassign
headlines_tokenized <- tib
```

\newpage

# Part 6

**Explore your data a bit and try to replicate some of the analyses above presented in class if you’d like (not necessary).**

\newpage

# Part 7

**Plot the amount of emotion words (the 8 from nrc) as a percentage of all the emotion words used each day (aggregate text from articles published on the same day). How does the distribution of emotion words change over time? Can you think of any reason this would be the case?**

```{r}
nrc_sent <- get_sentiments('nrc')

lunch_sent_counts <- 
```

\newpage