---
title: "EDS 231: Assignment 3"
author: "Scout Leonard"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: pdf_document
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

# Load Libraries

The code chunk below loads libraries used in this homework's R code.

```{r packages, results='hide', message=FALSE, warning=FALSE}
packages=c("quanteda.sentiment",
           "quanteda.textstats",
           "tidyverse",
           "tidytext",
           "lubridate",
           "here",
           "wordcloud", #visualization of common words in the data set
           "reshape2",
           "knitr",
           "reshape",
           "patchwork",
           "MetBrewer",
           "dplyr",
           "sentimentr",
           "quanteda") 

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}
```

# Import Data

\noindent The code chunk below calls data from the GitHub of my EDS 242 instructor. The data is Twitter data that includes tweets about International Panel on Climate Change (IPCC) from select dates before and after the publication of the 2022 IPCC report. The chunk also does some cleaning for the first parts of the homework. `raw_tweets` is used throughout.

```{r read in data }
raw_tweets <- read.csv("https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/main/dat/IPCC_tweets_April1-10_sample.csv", header=TRUE) 

dat<- raw_tweets[,c(4,6)] # Extract Date and Title fields

tweets <- tibble(text = dat$Title,
                  id = seq(1:length(dat$Title)),
                 date = as.Date(dat$Date,'%m/%d/%y'))
```

\newpage

# Part 1

\noindent Think about how to further clean a twitter data set. Let's assume that the mentions of twitter accounts is not useful to us. Remove them from the text field of the tweets tibble.

## Remove mentions and clean

\noindent The code chunk below cleans data containing tweet content by removing mentions, stop words, numbers, and meaningless URL bits so that the text analysis can focus on words meaningful to analyzing IPCC text and sentiment.

```{r tokenize + clean words from tweets}
#remove mentions
tweets$text <- str_remove(tweets$text, "@[a-z,A-Z]*")
tweets$text <- str_remove(tweets$text, "[:digit:]")

#tokenise tweets and remove stop words
words <- tweets %>%
  select(id, date, text) %>%
  unnest_tokens(output = word, input = text, token = "words") %>%
  anti_join(stop_words, by = "word")

#clean tokens
##remove numbers
clean_tokens <- str_remove_all(words$word, "[:digit:]")
##remove mentions
clean_tokens <- str_remove_all(clean_tokens, "@[a-z,A-Z]*")
##removes appostrophes
clean_tokens <- gsub("’s", '', clean_tokens)
## remove weird twitter and internet things
clean_tokens <- str_remove_all(clean_tokens, "https")
clean_tokens <- str_remove_all(clean_tokens, "t.co")

words$clean <- clean_tokens

#remove the empty strings
tib <-subset(words, clean != "")

#reassign
words <- tib
```

\newpage

# Part 2

\noindent Compare the ten most common terms in the tweets per day. Do you notice anything interesting?

## Count word use frequency and filter data for the top 10

```{r top 10 words}
#select the top 10 used words
words_freq <- words %>% 
  group_by(clean) %>% 
  summarise(freq = n()) %>% 
   top_n(10) %>% 
  select(clean)

#join with full dataset to return frquency of those words / day
top_10_words <- inner_join(words_freq, words, by = "clean") %>% 
  group_by(date, clean) %>% 
  summarize(freq = n()) 
```

## Plot use of top 10 words over time

\noindent The code chunk below plots a time series of how the top 10 words in IPCC tweet data changes over time - before, during, and after publication of the 2022 report.

```{r plot top 10 words}
ggplot(data = top_10_words, aes(x = date, y = freq)) +
  geom_line(aes(color = clean)) +
  theme_minimal() +
  labs(title = "Top 10 Most Frequent Words in IPCC-related Tweets",
       subtitle = "Tweets Leading Up to the 2022 Report",
       x = "Date",
       y = "Frequency",
       color = "Word") +
  theme(plot.title = element_text(hjust = 0.7, size = 15),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_color_manual(values = met.brewer("Nizami", n = 10))
```

\noindent I notice that there is a spike in all the top 10 words on April 4th, when the report is published. I assume this is because there is lots of coverage on this day compared to others.

\noindent "IPCC", "report", and "climate" are all the most frequent words the day the report is published and tend to stay highest for several days after, even as the frequency of all words declines.

\newpage

# Part 3

\noindent Adjust the wordcloud in the "wordcloud" chunk by coloring the positive and negative words so they are identifiable.

## Plot wordcloud

```{r plot colored wordcloud, fig.width= 4, fig.height=4, warning = FALSE, message = FALSE}
words %>%
inner_join(get_sentiments("bing")) %>%
count(word, sentiment, sort = TRUE) %>%
acast(word ~ sentiment, value.var = "n", fill = 0) %>%
comparison.cloud(colors = c("darkslateblue", "goldenrod"),
                   max.words = 100)
```

\newpage

# Part 4

\noindent Let's say we are interested in the most prominent entities in the Twitter discussion. Which are the top 10 most tagged accounts in the data set. Hint: the "explore_hashtags" chunk is a good starting point.

\noindent The code chunk below tokenizes words from tweets and removes puctuation except for the "@" symbol, so that I can count the most frequntly metioned Twitter accounts in the IPCC tweets. The 10 most frequent accounts are printed in the table below. 

```{r top 10 accounts, message = FALSE}
corpus <- corpus(dat$Title)

#tokenize words from Tweets
tokens <- tokens(corpus)

#clean up tokens
tokens <- tokens(tokens, remove_punct = TRUE,
                      remove_numbers = TRUE)

tokens <- tokens_select(tokens, stopwords('english'),selection='remove') #stopwords lexicon built in to quanteda

tokens <- tokens_wordstem(tokens) 

tokens <- tokens_tolower(tokens)

at_tweets <- tokens(corpus, remove_punct = TRUE) %>% 
               tokens_keep(pattern = "@*")

dfm_at<- dfm(at_tweets) #document feature matrix

tstat_freq <- textstat_frequency(dfm_at, n = 100)

#tidytext gives us tools to convert to tidy from non-tidy formats
at_tib<- tidy(dfm_at)

#print table of 10 most mentioned accounts
at_tib %>%
   group_by(term) %>% 
  summarise(n()) %>% 
  top_n(10) %>% 
  kable()
```

\newpage

# Part 5

The Twitter data download comes with a variable called "Sentiment" that must be calculated by Brandwatch. Use your own method to assign each tweet a polarity score (Positive, Negative, Neutral) and compare your classification to Brandwatch's (hint: you'll need to revisit the "raw_tweets" data frame).

## Use my method to assign sentiment

\noindent In the code chunk below, I get the sentiment of the raw tweets and add those sentiments to the raw tweets dataframe:

```{r get sentiment of raw tweets, warning = FALSE}
#give raw tweets element id
raw_tweets_trimmed <- data.frame(element_id = seq(1:length(raw_tweets$Title)), 
                       Date = raw_tweets$Date, 
                       Title = raw_tweets$Title)

#make a character vector to get sentiments from
tweets_text <- raw_tweets_trimmed$Title

#calculate the mean sentiment based on all sentences sentences in single tweet 
tweet_sent <- sentiment(tweets_text) %>% 
  group_by(element_id) %>% 
  summarise(mean_sent = mean(sentiment))

#join with sentence data
tweet_sent <- inner_join(raw_tweets_trimmed, tweet_sent, by = "element_id")
```

\noindent Next, I categorize the mean sentiment value for all of the sentences in a tweet as positive, negative, or neutral. I add these categories to my original dataframe of raw tweets so that I can compare my categorization method with the Brandwatch classification results. 

```{r categorize tweets}
# catgorize for ggplot
tweet_cat <- tweet_sent %>%  
  mutate(sent_cat = case_when(
    mean_sent < 0 ~ "negative",
    mean_sent == 0 ~ "neutral",
    mean_sent > 0 ~ "positive"
  ))

raw_tweets <- raw_tweets %>%
  mutate(element_id = seq(1:length(Title)))

all_sent <- full_join(tweet_cat, raw_tweets, by = "element_id", copy = TRUE) %>% 
  mutate_all(na_if, "")
```

## Plot both methods and compare

\noindent Finally, I plot the frequency of IPCC tweets classified as positive, negative, or neutral using the two methods, mine and Brandwatch:

```{r compare sentiment analyses, fig.width = 6, fig.height = 4}
p1 <- ggplot(data = all_sent, aes(x = sent_cat)) +
  geom_bar(stat = "count", 
           aes(fill = sent_cat),
           show.legend = FALSE) +
  theme_classic() +
  labs(title = "Scout's Classification of IPCC Tweet Sentiment",
       x = "Sentiment",
       y = "Frequency of Tweet Sentiment") +
  theme(plot.title = element_text(size = 12),
        axis.title.y = element_text(size = 7)) +
  scale_fill_manual(values = met.brewer("Degas", 3))

p2 <- ggplot(data = all_sent, aes(x = Sentiment)) +
  geom_bar(stat = "count",
           aes(fill = Sentiment),
           show.legend = FALSE) +
  theme_classic() +
  theme(plot.title = element_text(size = 12),
        axis.title.y = element_text(size = 7)) +
  labs(title = "Brandwatch Classification of IPCC Tweet Sentiment",
       y = "Frequency of Tweet Sentiment") +
  scale_fill_manual(values = met.brewer("Degas", n = 3))

p1 / p2 
```

\noindent The resulting figure shows that the Brandwatch sentiment analysis classification of tweet sentiment assigned many more tweets as "neutral" compared to my method, which assigns  tweets as negative or positive in a less conservative way. 
