---
title: 'Week 5: Assignment 4: Word relationship analysis'
author: "Scout Leonard"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries

```{r packages, results='hide', message=FALSE, warning=FALSE}
packages=c("tidyr",
           "pdftools",
           "lubridate",
           "tidyverse",
           "tidytext",
           "readr",
           "quanteda",
           "readtext",
           "quanteda.textstats",
           "quanteda.textplots",
           "ggplot2",
           "forcats",
           "stringr",
           "quanteda.textplots",
           "widyr",
           "igraph",
           "ggraph",
           "here") 

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}
```

# Read in data

```{r pdf_import}
#filepath to data
files <- list.files(path = here("data/week5_data/"), 
                    pattern = "*pdf$", 
                    full.names = TRUE)

#renders all textboxes on a text canvas and returns a character vector of equal length to the number of pages in the PDF file - a list of text from each file
ej_reports <- lapply(files, pdf_text)

#read texts and (if any) associated document-level meta-data from one or more source files - makes a df with each pdf and its text as a var
ej_pdf <- readtext(file = here("data/week5_data/*.pdf"), 
                   docvarsfrom = "filenames",
                   docvarnames = c("type", "subj", "year"),
                   sep = "_")

#creating an initial corpus containing our data
epa_corp <- corpus(x = ej_pdf, text_field = "text" )

#return details of the corpus
summary(epa_corp) %>% 
  knitr::kable()

#I'm adding some additional, context-specific stop words to stop word lexicon
more_stops <-c("2015","2016", "2017", "2018",
               "2019", "2020", "www.epa.gov", "https")

#add the additional stopwords to the stop word lexicon
add_stops <- tibble(word = c(stop_words$word, more_stops)) 

stop_vec <- as_vector(add_stops)
```

```{r tidy}
#convert to tidy format and apply my stop words
raw_text <- tidy(epa_corp)

#Distribution of most frequent words across documents
raw_words <- raw_text %>%
  mutate(year = as.factor(year)) %>%
  unnest_tokens(word, text) %>%
  anti_join(add_stops, by = 'word') %>%
  count(year, word, sort = TRUE)

#number of total words by document  
total_words <- raw_words %>% 
  group_by(year) %>% 
  summarize(total = sum(n))

report_words <- left_join(raw_words, total_words)
 
par_tokens <- unnest_tokens(raw_text, 
                            output = paragraphs, 
                            input = text, 
                            token = "paragraphs")

par_tokens <- par_tokens %>%
 mutate(par_id = 1:n())

par_words <- unnest_tokens(par_tokens,
                           output = word,
                           input = paragraphs,
                           token = "words")
```

\newpage

# Part 1

\noindent What are the most frequent trigrams in the dataset? How does this compare to the most frequent bigrams? Which n-gram seems more informative here, and why?

```{r quanteda_init}
tokens <- tokens(epa_corp, remove_punct = TRUE) #list of character vectors - takes each document and splits it word by word

toks1<- tokens_select(tokens, min_nchar = 3)

toks1 <- tokens_tolower(toks1)

toks1 <- tokens_remove(toks1, pattern = (stop_vec))

dfm <- dfm(toks1) #create document feature matrix - rows are number of occurances of each word within each document 

#first the basic frequency stat
tstat_freq <- textstat_frequency(dfm, n = 5, groups = year)

head(tstat_freq, 10) %>% 
  knitr::kable()
```


```{r convert_dfm}
toks2 <- tokens_ngrams(toks1, n = 3)

dfm2 <- dfm(toks2)

dfm2 <- dfm_remove(dfm2, pattern = c(stop_vec))
#gives more coherent terms - power of chunking at a different token level

freq_words2 <- textstat_frequency(dfm2, n = 20)

freq_words2$token <- rep("trigram", 20)
#tokens1 <- tokens_select(tokens1,pattern = stopwords("en"), selection = "remove")

head(freq_words2, 5) %>% 
  knitr::kable()
```

\noindent The most frequent trigrams in the dataset are shown in the table above, with `r freq_words2$feature[1]` as the most frequently occurring trigram. 

\newpage

# Part 2

\noindent Choose a new focal term to replace "justice" and recreate the correlation table and network (see corr_paragraphs and corr_network chunks). Explore some of the plotting parameters in the cor_network chunk to see if you can improve the clarity or amount of information your plot conveys. Make sure to use a different color for the ties!

```{r corr_paragraphs, warning = FALSE}
#correlation between co-occuring words
word_cors <- par_words %>% 
  add_count(par_id) %>% 
  filter(n >= 50) %>% 
  select(-n) %>%
  pairwise_cor(word, par_id, sort = TRUE)

#now we can select words cooccurring with the word justice and get correlation coefficients 
food_cors <- word_cors %>% 
  filter(item1 == "food")

  word_cors %>%
  filter(item1 %in% c("food", "agriculture", "health", "resiliency")) %>%
  group_by(item1) %>%
  top_n(6) %>%
  ungroup() %>%
  mutate(item1 = as.factor(item1),
  name = reorder_within(item2, correlation, item1)) %>%
  ggplot(aes(y = name, x = correlation, fill = item1)) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~item1, ncol = 2, scales = "free")+
  scale_y_reordered() +
  labs(y = NULL,
         x = NULL,
         title = "Correlations with key words",
         subtitle = "EPA EJ Reports")
  
  #let's zoom in on just one of our key terms
   food_cors <- word_cors %>% 
  filter(item1 == "food") %>%
   mutate(n = 1:n())
```

```{r corr_network, warning = FALSE}
food_cors  %>%
  filter(n <= 50) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation,
                     edge_width = correlation),
                 edge_colour = "darkmagenta") +
  geom_node_point(size = 3) +
  geom_node_text(aes(label = name),
                 repel = TRUE, 
                 point.padding = unit(0.2,
                                      "lines")) +
  theme_void()
```

\newpage

# Part 3

\noindent Write a function that allows you to conduct a keyness analysis to compare two individual EPA reports (hint: that means target and reference need to both be individual reports). Run the function on 3 pairs of reports, generating 3 keyness plots.

```{r}
#test function
dfm_subset <- corpus_subset(epa_corp, grepl("2018|2019", docnames(epa_corp)))
```


```{r}
#write function
keyness_comparison <- function(text1_year, text2_year) {
  
  #subset the corpus
  corpus_subset <- corpus_subset(epa_corp, grepl(paste0(text1_year, "|", text2_year), docnames(epa_corp)))
  
  #tokenize corpus
  tokens <- tokens(corpus_subset, remove_punct = TRUE) #list of character vectors - takes each document and splits it word by word
  toks<- tokens_select(tokens, min_nchar = 3)
  toks <- tokens_tolower(toks)
  toks <- tokens_remove(toks, pattern = (stop_vec))
  dfm <- dfm(toks) #create document feature matrix - rows are number of occurances of each word within each document 
  
  keyness <- textstat_keyness(dfm, target = 2)
  
  textplot_keyness(keyness)
  
}

```

\newpage

## Keyness plot 1

Test running the function on 2015 and 2019

```{r}
keyness_comparison(text1_year = 2015, text2_year = 2019)
```

\newpage

## Keyness plot 2

Test running the function on 2019 and 2020

```{r}
keyness_comparison(text1_year = 2019, text2_year = 2020)
```

\newpage

## Keyness plot 3

Test running the function on 2016 and 2018:

```{r}
keyness_comparison(text1_year = 2016, text2_year = 2018)
```

\newpage

# Part 4

\noindent Select a word or multi-word term of interest and identify words related to it using windowing and keyness comparison. To do this you will create two objects: one containing all words occurring within a 10-word window of your term of interest, and the second object containing all other words. Then run a keyness comparison on these objects. Which one is the target, and which the reference? [Hint](https://tutorials.quanteda.io/advanced-operations/target-word-collocations/)

**food systems**

Create an object containing all words occurring within a 10 word window of **food systems**. 

```{r}
term = "food"

toks_inside <- tokens_keep(toks1,
                           pattern = term,
                           window = 10)

toks_inside <- tokens_remove(toks_inside,
                             pattern = term)
```

Create an object containing all other words: 

```{r}
toks_outside <- tokens_remove(toks1,
                              pattern = term,
                              window = 10)
```

Run a keyness comparison of the objects: 

```{r}
dfmat_inside <- dfm(toks_inside)
dfmat_outside <- dfm(toks_outside)

tstat_key_inside <- textstat_keyness(rbind(dfmat_inside, dfmat_outside), 
                                     target = seq_len(ndoc(dfmat_inside)))

textplot_keyness(tstat_key_inside)
```

`toks_inside` is the target, and `toks_outside` is the reference. 

\newpage