---
title: 'EDS 231: Week 4 lab: Sentiment Analysis II'
author: "Scout Leonard"
date: '2022-04-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following Rmarkdown workflow was completed as a lab for my course EDS 231: Text and Sentiment Analysis for Environmental Data Science. [See the course website for more information.](https://maro406.github.io/EDS_231-text-sentiment/topic_4.html)

#### IPCC Report Twitter

```{r packages, results='hide', message=FALSE, warning=FALSE}
packages=c("quanteda.sentiment",
           "quanteda.textstats",
           "tidyverse",
           "tidytext",
           "lubridate",
           "here",
           "wordcloud", #visualization of common words in the data set
           "reshape2",
           "quanteda") #devtools::install_github("quanteda/quanteda.sentiment") #not available currently through CRAN

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}
```

Last week we used the tidytext approach to sentiment analysis for Nexis Uni .pdf data on coverage of the recent IPCC report. This week we will look at the conversation on Twitter about the same report. We'll start with the familiar tidy approach, and then introduce the quanteda package later.

```{r tweet_data}

raw_tweets <- read.csv("https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/main/dat/IPCC_tweets_April1-10_sample.csv", header=TRUE)

dat<- raw_tweets[,c(5,7)] # Extract Date and Title fields

tweets <- tibble(text = dat$Title,
                  id = seq(1:length(dat$Title)),
                 date = as.Date(dat$Date,'%m/%d/%y'))


head(tweets$text, n = 10)

#simple plot of tweets per day
tweets %>%
  count(date) %>%
  ggplot(aes(x = date, y = n))+
  geom_line()

```

```{r cleaning_tweets}

#let's clean up the URLs from the tweets
tweets$text <- gsub("http[^[:space:]]*", "",tweets$text)
tweets$text <- str_to_lower(tweets$text)

#load sentiment lexicons
bing_sent <- get_sentiments('bing')
nrc_sent <- get_sentiments('nrc')

#tokenize tweets to individual words
words <- tweets %>%
  select(id, date, text) %>%
  unnest_tokens(output = word, input = text, token = "words") %>%
  anti_join(stop_words, by = "word") %>%
  left_join(bing_sent, by = "word") %>%
  left_join(
    tribble(
      ~sentiment, ~sent_score,
      "positive", 1,
      "negative", -1),
    by = "sentiment")
```

What is the sentiment associated with the things people are saying about this particular topic? 

```{r sentiment_calculations}
#take average sentiment score by tweet
tweets_sent <- tweets %>%
  left_join(
    words %>%
      group_by(id) %>%
      summarize(
        sent_score = mean(sent_score, na.rm = T)),
    by = "id")

neutral <- length(which(tweets_sent$sent_score == 0))
positive <- length(which(tweets_sent$sent_score > 0))
negative <- length(which(tweets_sent$sent_score < 0))

Sentiment <- c("Positive","Neutral","Negative")
Count <- c(positive,neutral,negative)
output <- data.frame(Sentiment,Count)
output$Sentiment<-factor(output$Sentiment,levels=Sentiment)
ggplot(output, aes(x=Sentiment,y=Count))+
  geom_bar(stat = "identity", aes(fill = Sentiment))+
  scale_fill_manual("legend", values = c("Positive" = "green", "Neutral" = "black", "Negative" = "red"))+
  ggtitle("Barplot of Sentiment in IPCC tweets")
```

Plot the average sentiment score and how it changes by day: 

```{r plot_sentiment_by_day}
# tally sentiment score per day
daily_sent <- tweets_sent %>%
  group_by(date) %>%
  summarize(sent_score = mean(sent_score, na.rm = T))

daily_sent %>%
  ggplot( aes(x = date, y = sent_score)) +
  geom_line() +
    labs(x = "Date",
    y = "Avg Sentiment Score",
    title = "Daily Tweet Sentiment",
    subtitle = "IPCC Tweets")

```

_Anticipation prior the the report boosts sentiment prior to it's release, than reactions bring it down after publication._

Now let's try a new type of text visualization: the wordcloud.

```{r wordcloud}
words %>%
   anti_join(stop_words) %>%
   count(word) %>%
   with(wordcloud(word, n, max.words = 100))


```

We can make it more informative by bring in our sentiment information:

```{r wordcloud_comp}

words %>%
inner_join(get_sentiments("bing")) %>%
count(word, sentiment, sort = TRUE) %>%
acast(word ~ sentiment, value.var = "n", fill = 0) %>%
comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

_This is a compartison word cloud, with positive words on bottom in grey and positive words on top in black._ 

#### The quanteda package

quanteda is a package (actually a family of packages) full of tools for conducting text analysis. quanteda.sentiment (not yet on CRAN, download from github) is the quanteda modular package for conducting sentiment analysis.

quanteda has its own built in functions for cleaning text data. Let's take a look at some. First we have to clean the messy tweet data:

_Corpus is a collection of documents and associated metadata._ 

```{r create_corpus}
corpus <- corpus(dat$Title) #enter quanteda
summary(corpus)
```

_Summary returns that corpus has 2400 documents. Each row is a tweet, detects sentences using puntuation, types refers to specific words and tokens refers to total words. You can have multiple words of the same type, so tokens will alwasys be great than type._

```{r quanteda_cleaning}
tokens <- tokens(corpus) #tokenize the text so each doc (page, in this case) is a list of tokens (words)

#examine the uncleaned version
tokens

#clean it up
tokens <- tokens(tokens, remove_punct = TRUE,
                      remove_numbers = TRUE)

tokens <- tokens_select(tokens, stopwords('english'),selection='remove') #stopwords lexicon built in to quanteda

#tokens <- tokens_wordstem(tokens) #stem words down to their base form for comparisons across tense and quantity

tokens <- tokens_tolower(tokens)

```

We can use the kwic function (keywords-in-context) to briefly examine the context in which certain words or patterns appear.

_kwic stands for key words in context. It goes through the corpus and pulls out all the words of a given pattern and returns the context._

```{r initial_analysis}
head(kwic(tokens, pattern = "climate", window = 3))

head(kwic(tokens, pattern = phrase("climate change"), window = 3))


```

_Anlayzing hastags present in tweets:_

_We take the corpus and keep only tokens with a certain apttern inside the corpus_

**Document feature matrix:** a matrix that show the location of each of the features or words in the corpus located in each of the tweets were they occur. This is the beginning of **word embeddings** or semantic understanding. If you want to know how similar words are, you can compare a vector of occurrences. 

```{r explore_hashtags}
hash_tweets <- tokens(corpus, remove_punct = TRUE) %>% 
               tokens_keep(pattern = "#*")

dfm_hash<- dfm(hash_tweets) #document feature matrix

tstat_freq <- textstat_frequency(dfm_hash, n = 100)
head(tstat_freq, 10)

#tidytext gives us tools to convert to tidy from non-tidy formats
hash_tib<- tidy(dfm_hash)

hash_tib %>%
   count(term) %>%
   with(wordcloud(term, n, max.words = 100))
```

Create the sparse matrix representation known as the document-feature matrix. quanteda's textstat_polarity function has multiple ways to combine polarity to a single score. The sent_logit value to fun argument is the log of (pos/neg) counts.

```{r}
dfm <- dfm(tokens)

topfeatures(dfm, 12)

dfm.sentiment <- dfm_lookup(dfm, dictionary = data_dictionary_LSD2015)

head(textstat_polarity(tokens, data_dictionary_LSD2015, fun = sent_logit))
```

### Assignment

You will use the tweet data from class today for each part of the following assignment.

1.  Think about how to further clean a twitter data set. Let's assume that the mentions of twitter accounts is not useful to us. Remove them from the text field of the tweets tibble.

2.  Compare the ten most common terms in the tweets per day.  Do you notice anything interesting?

3.  Adjust the wordcloud in the "wordcloud" chunk by coloring the positive and negative words so they are identifiable.

4. Let's say we are interested in the most prominent entities in the Twitter discussion.  Which are the top 10 most tagged accounts in the data set. Hint: the "explore_hashtags" chunk is a good starting point.

5. The Twitter data download comes with a variable called "Sentiment" that must be calculated by Brandwatch.  Use your own method to assign each tweet a polarity score (Positive, Negative, Neutral) and compare your classification to Brandwatch's (hint: you'll need to revisit the "raw_tweets" data frame).   
