---
title: 'Week 2 Lab: Text Data in R'
author: "Scout Leonard"
date: "4/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Connect to the New York Times API and send a query

I first needed to create an API key after making a [NYT delveloper account.](https://developer.nytimes.com/). 

## Load Libraries

```{r}
library(jsonlite) #convert results from API queries into R-friendly formats 
library(tidyverse) 
library(tidytext) #text data management and analysis
library(ggplot2) #plot word frequencies and publication dates
library(here)
```

This connects to the article search API form the New York Times. 

```{r}
#create an object called t with the results of our query ("haaland")
# the from JSON flatten the JSON object, then convert to a data frame
t <- fromJSON("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=haaland&api-key=NTKBHbsb6XFEkGymGumAiba7n3uBvs8V", flatten = TRUE) #the string following "key=" is your API key 

class(t) #what type of object is t?

t <- t %>% 
  data.frame()


#Inspect our data
class(t) #now what is it?
dim(t) # how big is it?
names(t) # what variables are we working with?
#t <- readRDS("nytDat.rds") #in case of API emergency :)
```

We have 10 articles and 33 fields/variables within each of those article objects. 

`names(t)` returns the 

```{r}
t$response.docs.snippet[9]

#assign a snippet to x to use as fodder for stringr functions.  You can follow along using the sentence on the next line.

x <- "Her nomination as secretary of the interior is historic, but as the first Native cabinet member, she would have to strike a delicate balance." 

tolower(x)
str_split(x, ','); str_split(x, 't') #splits by commas, then by the letter t
str_replace(x, 'historic', 'without precedent')
str_replace(x, ' ', '_') #first one
#how do we replace all of them?
str_replace_all(x, ' ', '_')
str_detect(x, 't'); str_detect(x, 'tive') ### is pattern in the string? T/F
str_locate(x, 't'); str_locate_all(x, 'as')
```

OK, it’s working but we want more data. Let’s set some parameters for a bigger query

```{r}
term <- "Haaland" # Need to use + to string together separate words
begin_date <- "20210120"
end_date <- "20220401"

#construct the query url using API operators
baseurl <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",term,
                  "&begin_date=",begin_date,"&end_date=",end_date,
                  "&facet_filter=true&api-key=","NTKBHbsb6XFEkGymGumAiba7n3uBvs8V", sep="")
```

```{r}
#examine our query url
#this code allows for obtaining multiple pages of query results 
#  initialQuery <- fromJSON(baseurl)
# maxPages <- round((initialQuery$response$meta$hits[1] / 10)-1) 
# 
# pages <- list()
# for(i in 0:maxPages){
#   nytSearch <- fromJSON(paste0(baseurl, "&page=", i), flatten = TRUE) %>% data.frame() 
#   message("Retrieving page ", i)
#   pages[[i+1]] <- nytSearch 
#   Sys.sleep(6) 
# }
# class(nytSearch)

#need to bind the pages and create a tibble from nytDa
```

I have to do some troubleshooting with my API, so I downloaded [a csv of the data from the instructor's Github.](https://github.com/MaRo406/EDS_231-text-sentiment/blob/main/nytDat.csv)

```{r}
nytDat <- read.csv(here("data","nytDat.csv")) # obtained from 

nytDat %>% 
  group_by(response.docs.type_of_material) %>%
  summarize(count=n()) %>%
  mutate(percent = (count / sum(count))*100) %>%
  ggplot() +
  geom_bar(aes(y=percent, x=response.docs.type_of_material, fill=response.docs.type_of_material), stat = "identity") + coord_flip()
```

```{r}
nytDat %>%
  mutate(pubDay=gsub("T.*","",response.docs.pub_date)) %>%
  group_by(pubDay) %>%
  summarise(count=n()) %>%
  filter(count >= 2) %>%
  ggplot() +
  geom_bar(aes(x=reorder(pubDay, count), y=count), stat="identity") + coord_flip()

```

```{r}
names(nytDat)
```

The code chunk below makes a bar graph that includes all of the words, so it is hard to distill any important messages about our topic of interest. We can manipulate this in the `filter()` argument . There are a lot of words that are common and don't give us much meaning: **in, and, a, of, that, etc.**

```{r}
paragraph <- names(nytDat)[6] #The 6th column, "response.doc.lead_paragraph", is the one we want here.  
tokenized <- nytDat %>%
  unnest_tokens(word, paragraph)

tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed at 5
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```
`stop_words` is a list that contains common stop words in the English language. 

```{r}
data(stop_words)

tokenized <- tokenized %>%
  anti_join(stop_words) #removes the stop words 

tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 5) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

```{r}
#inspect the list of tokens (words)
tokenized$word

clean_tokens <- str_replace_all(tokenized$word,"land[a-z,A-Z]*","land") #stem tribe words
clean_tokens <- str_remove_all(clean_tokens, "[:digit:]") #remove all numbers
clean_tokens <- str_remove_all(clean_tokens, "washington")

clean_tokens <- gsub("’s", '', clean_tokens) #global substitution - base r function

tokenized$clean <- clean_tokens

tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)

#remove the empty strings
tib <-subset(tokenized, clean!="")

#reassign
tokenized <- tib

#try again
tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)
```

