---
title: "EDS 231: Assignment 6: Word Embeddings"
author: "Scout Leonard"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: pdf_document
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

Download a set of pretrained vectors, GloVe, and explore them.

Grab data here: <!-- download.file('<https://nlp.stanford.edu/data/glove.6B.zip>',destfile = 'glove.6B.zip')  --> <!-- unzip('glove.6B.zip')  --> <!-- Use this file: 'glove.6B.300d.txt' -->

# Load Libraries

```{r}

```


# Download Data 

```{r}

```

# Assignment

_Download a set of pretrained vectors, GloVe, and explore them._ 

1. Recreate the analyses in the last three chunks (find-synonyms, plot-synonyms, word-math) with the GloVe embeddings. How are they different from the embeddings created from the climbing accident data? Why do you think they are different?

2. Run the classic word math equation, "king" - "man" = ?

3. Think of three new word math equations. They can involve any words you'd like, whatever catches your interest.